\textbf{The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS}

### 2025-08-12 paper/manuscript.qmd
boom

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Emma Benn 
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Maisie Bennett
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  TO DO.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness, Health"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

<!-- There are many questionnaire measures – 14 of which are commonly used and 5 (BPQ, MAIA, BAQ, SAQ, and PBCS) which cover 86.5% of all citations of interoceptive sensibility questionnaires – Desmedt et al., 2022 -->
The present work focuses on self-report questionnaires, which offer a scalable and accessible tool complementing physiological tasks. Among the numerous questionnaires available, a handful seem to dominate the literature [@desmedt2022measures], including the Body Perception Questionnaire [BPQ, @kolacz2018body], the Multidimensional Assessment of Interoceptive Awareness [MAIA, @mehling2018multidimensional], and the Interoceptive Accuracy Scale [IAS, @murphy2018classifying]. These tools were developed in different eras of a rapidly evolving field, each with its own objectives and philosophies. The BPQ, one of the earliest measures, emphasizes the assessment of body awareness and autonomic nervous system (ANS) reactivity, focusing on subjective experiences of physiological stress reactions in organs innervated by the ANS, such as those triggered by stress or trauma [@porges1993body]. The MAIA incorporated ideas from the then-booming emotion regulation and mindfulness fields, including subscales assessing adaptive coping strategies and metacognitive insights about the body. By contrast, the IAS was recently crafted to sidestep such high-level - possibly tangential - constructs and instead capture subjective reports of concrete bodily events (e.g., coughing, sneezing) as proxies for interoceptive accuracy.

@@

While the overall structure of the Mint is stable, some facets warrant deeper investigation. The *Olfa* (Olfactory Compensation) and *Derm* (Dermal Hypersensitivity) clusters, for instance, showed more complex patterns of association. Their placement within the *Deficit* metacluster is theoretically intriguing but requires further validation. The former aligns with predictive coding models suggesting that individuals may rely on exteroceptive cues (like odours) to infer internal states  [@paulus2010interoception], a form of pseudo-interoception (or extero-interoception). The latter points to a potential link between visceral and cutaneous sensitivity, possibly reflecting a general trait of heightened bodily awareness or a shared mechanism of central sensitization, which has been observed in conditions like chronic pain [@di2016alexithymia]. Future studies should examine the invariance of the Mint's structure across different populations (e.g., clinical vs. non-clinical groups) and contexts (e.g., heightened arousal vs. rest) to clarify the role of these more nuanced facets. Such work could reveal whether their relevance and pattern of associations differs in conditions like anxiety disorders, chronic pain, or neurodevelopmental disorders.

>>>>

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Emma Benn 
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Maisie Bennett
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  Interoception, the sense of the body's internal state, is crucial for mental and physical health, yet its measurement is hindered by tools with conceptual and psychometric issues. To address these limitations, we developed and validated the Multimodal Interoception Questionnaire (Mint). In Study 1 (N=559), we generated a comprehensive item pool using a novel "modality-by-context" framework and employed the recently developed psychometric network analysis for data-driven item reduction. In Study 2 (N=737), we validated the resulting instrument in an independent sample, establishing its structure and comparing its predictive power against established interoception questionnaires (MAIA-2, BPQ, IAS) across a wide range of clinical and cognitive outcomes. This rigorous process yielded a 33-item scale with a stable hierarchical structure comprising three metaclusters (Interoceptive Deficit, Interoceptive Awareness, Visceroception) and 11 distinct facets. The Mint demonstrated superior criterion validity, consistently outperforming existing scales in predicting outcomes tied to altered bodily perception, including alexithymia, ADHD, autism, and somatic symptom clusters. The Mint offers a psychometrically robust and nuanced measure of self-reported interoception that integrates and extends existing scales, providing researchers and clinicians with a practical and readily usable tool to investigate the role of bodily sensations in health and disease.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness, Health"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

<!-- There are many questionnaire measures – 14 of which are commonly used and 5 (BPQ, MAIA, BAQ, SAQ, and PBCS) which cover 86.5% of all citations of interoceptive sensibility questionnaires – Desmedt et al., 2022 -->
The present work focuses on self-report questionnaires, which offer a scalable and accessible tool complementing physiological tasks. Among the numerous questionnaires available, a handful seem to dominate the literature [@desmedt2022measures], including the Body Perception Questionnaire [BPQ, @kolacz2018body], the Multidimensional Assessment of Interoceptive Awareness [MAIA, @mehling2018multidimensional], and the Interoceptive Accuracy Scale [IAS, @murphy2018classifying]. These tools were developed in different eras of a rapidly evolving field, each with its own objectives and philosophies. The BPQ, one of the earliest measures, emphasizes the assessment of body awareness and autonomic nervous system (ANS) reactivity, focusing on subjective experiences of physiological stress reactions in organs innervated by the ANS, such as those triggered by stress or trauma [@kolacz2018body]. The MAIA incorporated ideas from the then-booming emotion regulation and mindfulness fields, including subscales assessing adaptive coping strategies and metacognitive insights about the body. By contrast, the IAS was recently crafted to sidestep such high-level - possibly tangential - constructs and instead capture subjective reports of concrete bodily events (e.g., coughing, sneezing) as proxies for interoceptive accuracy.

@@

While the overall structure of the Mint is stable, some facets warrant deeper investigation. The *Olfa* (Olfactory Compensation) and *Derm* (Dermal Hypersensitivity) clusters, for instance, showed more complex patterns of association. Their placement within the *Deficit* metacluster is theoretically intriguing but requires further validation. The former aligns with predictive coding models suggesting that individuals may rely on exteroceptive cues (like odours) to infer internal states  [@paulus2010interoception], a form of pseudo-interoception (or extero-interoception). The latter points to a potential link between visceral and cutaneous sensitivity, possibly reflecting a general trait of heightened bodily awareness or a shared mechanism of central sensitization, which has been observed in conditions like chronic pain [@di2016alexithymia]. Future studies should examine the invariance of the Mint's structure across different populations (e.g., clinical vs. non-clinical groups) and contexts (e.g., heightened arousal vs. rest) to clarify the role of these more nuanced facets. Such work could reveal whether their relevance and pattern of associations differ in conditions like anxiety disorders, chronic pain, or neurodevelopmental disorders.

### 2025-08-12 paper/manuscript.tex
boom

\abstract{TO DO. }

@@

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, toprule=.15mm, arc=.35mm, left=2mm, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0, colback=white]

@@

The present work focuses on self-report questionnaires, which offer a
scalable and accessible tool complementing physiological tasks. Among
the numerous questionnaires available, a handful seem to dominate the
literature (\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}),
including the Body Perception Questionnaire (BPQ,
\citeproc{ref-kolacz2018body}{Kolacz et al., 2018}), the
Multidimensional Assessment of Interoceptive Awareness (MAIA,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}), and
the Interoceptive Accuracy Scale (IAS,
\citeproc{ref-murphy2018classifying}{Murphy et al., 2018}). These tools
were developed in different eras of a rapidly evolving field, each with
its own objectives and philosophies. The BPQ, one of the earliest
measures, emphasizes the assessment of body awareness and autonomic
nervous system (ANS) reactivity, focusing on subjective experiences of
physiological stress reactions in organs innervated by the ANS, such as
those triggered by stress or trauma
(\citeproc{ref-porges1993body}{\textbf{porges1993body?}}). The MAIA
incorporated ideas from the then-booming emotion regulation and
mindfulness fields, including subscales assessing adaptive coping
strategies and metacognitive insights about the body. By contrast, the
IAS was recently crafted to sidestep such high-level - possibly
tangential - constructs and instead capture subjective reports of
concrete bodily events (e.g., coughing, sneezing) as proxies for
interoceptive accuracy.

@@

Exploratory correlations with the emotion regulation (CERQ) dimensions
revealed stronger associations with most of the MAIA dimensions
(supporting its proximity with emotion regulation). Interestingly,
Rumination (and Self-Blame) stood out as selectively related to the
Mint's \emph{Visceroception} and \emph{Deficit} dimensions (note that
Rumination was also related to the MAIA's interoceptive dimensions,
namely Noticing, Emotional Awareness and Body Listening). The only
primal world belief that correlated particularly with the Mint's
\emph{Visceroception} and the MAIA's interoceptive dimension was the
belief that the world was alive (\textbf{TODO: Ana maybe throw in a
couple more words about what this belief means and is defined}).

@@

While the overall structure of the Mint is stable, some facets warrant
deeper investigation. The \emph{Olfa} (Olfactory Compensation) and
\emph{Derm} (Dermal Hypersensitivity) clusters, for instance, showed
more complex patterns of association. Their placement within the
\emph{Deficit} metacluster is theoretically intriguing but requires
further validation. The former aligns with predictive coding models
suggesting that individuals may rely on exteroceptive cues (like odours)
to infer internal states (\citeproc{ref-paulus2010interoception}{Paulus
\& Stein, 2010}), a form of pseudo-interoception (or
extero-interoception). The latter points to a potential link between
visceral and cutaneous sensitivity, possibly reflecting a general trait
of heightened bodily awareness or a shared mechanism of central
sensitization, which has been observed in conditions like chronic pain
(\citeproc{ref-di2016alexithymia}{Di Tella \& Castelli, 2016}). Future
studies should examine the invariance of the Mint's structure across
different populations (e.g., clinical vs.~non-clinical groups) and
contexts (e.g., heightened arousal vs.~rest) to clarify the role of
these more nuanced facets. Such work could reveal whether their
relevance and pattern of associations differs in conditions like anxiety
disorders, chronic pain, or neurodevelopmental disorders.

>>>>

\abstract{Interoception, the sense of the body's internal state, is
crucial for mental and physical health, yet its measurement is hindered
by tools with conceptual and psychometric issues. To address these
limitations, we developed and validated the Multimodal Interoception
Questionnaire (Mint). In Study 1 (N=559), we generated a comprehensive
item pool using a novel ``modality-by-context'' framework and employed
the recently developed psychometric network analysis for data-driven
item reduction. In Study 2 (N=737), we validated the resulting
instrument in an independent sample, establishing its structure and
comparing its predictive power against established interoception
questionnaires (MAIA-2, BPQ, IAS) across a wide range of clinical and
cognitive outcomes. This rigorous process yielded a 33-item scale with a
stable hierarchical structure comprising three metaclusters
(Interoceptive Deficit, Interoceptive Awareness, Visceroception) and 11
distinct facets. The Mint demonstrated superior criterion validity,
consistently outperforming existing scales in predicting outcomes tied
to altered bodily perception, including alexithymia, ADHD, autism, and
somatic symptom clusters. The Mint offers a psychometrically robust and
nuanced measure of self-reported interoception that integrates and
extends existing scales, providing researchers and clinicians with a
practical and readily usable tool to investigate the role of bodily
sensations in health and disease. }

@@

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, leftrule=.75mm, opacityback=0, colframe=quarto-callout-note-color-frame, breakable, left=2mm, arc=.35mm, colback=white, bottomrule=.15mm, toprule=.15mm]

@@

The present work focuses on self-report questionnaires, which offer a
scalable and accessible tool complementing physiological tasks. Among
the numerous questionnaires available, a handful seem to dominate the
literature (\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}),
including the Body Perception Questionnaire (BPQ,
\citeproc{ref-kolacz2018body}{Kolacz et al., 2018}), the
Multidimensional Assessment of Interoceptive Awareness (MAIA,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}), and
the Interoceptive Accuracy Scale (IAS,
\citeproc{ref-murphy2018classifying}{Murphy et al., 2018}). These tools
were developed in different eras of a rapidly evolving field, each with
its own objectives and philosophies. The BPQ, one of the earliest
measures, emphasizes the assessment of body awareness and autonomic
nervous system (ANS) reactivity, focusing on subjective experiences of
physiological stress reactions in organs innervated by the ANS, such as
those triggered by stress or trauma
(\citeproc{ref-kolacz2018body}{Kolacz et al., 2018}). The MAIA
incorporated ideas from the then-booming emotion regulation and
mindfulness fields, including subscales assessing adaptive coping
strategies and metacognitive insights about the body. By contrast, the
IAS was recently crafted to sidestep such high-level - possibly
tangential - constructs and instead capture subjective reports of
concrete bodily events (e.g., coughing, sneezing) as proxies for
interoceptive accuracy.

@@

Exploratory correlations with the emotion regulation (CERQ) dimensions
revealed stronger associations with most of the MAIA dimensions
(supporting its proximity with emotion regulation). Interestingly,
Rumination (and Self-Blame) stood out as selectively related to the
Mint's \emph{Visceroception} and \emph{Deficit} dimensions (note that
Rumination was also related to the MAIA's interoceptive dimensions,
namely Noticing, Emotional Awareness and Body Listening). The only
primal world belief that correlated particularly with the Mint's
\emph{Visceroception} and the MAIA's interoceptive dimension was the
belief that the world was alive (i.e., the belief that the universe is
conscious, purposeful, and in active relationship with oneself,
\citeproc{ref-clifton2021brief}{Clifton \& Yaden, 2021}).

@@

While the overall structure of the Mint is stable, some facets warrant
deeper investigation. The \emph{Olfa} (Olfactory Compensation) and
\emph{Derm} (Dermal Hypersensitivity) clusters, for instance, showed
more complex patterns of association. Their placement within the
\emph{Deficit} metacluster is theoretically intriguing but requires
further validation. The former aligns with predictive coding models
suggesting that individuals may rely on exteroceptive cues (like odours)
to infer internal states (\citeproc{ref-paulus2010interoception}{Paulus
\& Stein, 2010}), a form of pseudo-interoception (or
extero-interoception). The latter points to a potential link between
visceral and cutaneous sensitivity, possibly reflecting a general trait
of heightened bodily awareness or a shared mechanism of central
sensitization, which has been observed in conditions like chronic pain
(\citeproc{ref-di2016alexithymia}{Di Tella \& Castelli, 2016}). Future
studies should examine the invariance of the Mint's structure across
different populations (e.g., clinical vs.~non-clinical groups) and
contexts (e.g., heightened arousal vs.~rest) to clarify the role of
these more nuanced facets. Such work could reveal whether their
relevance and pattern of associations differ in conditions like anxiety
disorders, chronic pain, or neurodevelopmental disorders.
